---
title: "Exercises 1.9-1.15"
author: "Aleksandra Kilijańska"
date: "2023-04-18"
output: html_document
---
## 1.9 (Custom standard deviation function).
Write a function sd.n similar to the function var.n in Example 1.5 that will return the estimate σˆ (the
square root of σˆ2). Try this function on the temperature data of Example 1.1.

```{r}
sd.n = function(x) {
  n = length(x)
  if (n == 1) {
    stop("Need at least two data points")
  }
  m = mean(x)
  s = sqrt(sum((x - m)^2) / (n - 1))
  return(s)
}

temps = c(51.9, 51.8, 51.9, 53)
sd.n(temps)

```
## 1.10 (Euclidean norm function). 
Write a function norm that will compute
the Euclidean norm of a numeric vector. The Euclidean norm of a vector
x = (x1,...,xn) is... Use vectorized operations to compute the sum. Try this function on the vectors (0,0,0,1) and (2,5,2,4) to check that your function result is correct.

```{r}
norm = function(x){
 
  return(sqrt(sum(x^2)))
}

v1=c(0,0,0,1)
norm(v1)

v2=c(2,5,2,4)
norm(v2)
```

## 1.11 (Numerical integration). 
Use the curve function to display the
graph of the function f(x) = e−x2
/(1 +x2) on the interval 0 ≤ x ≤ 10. Then
use the integrate function to compute the value of the integral. The upper limit at infinity is specified by upper=Inf in the integrate function.
```{r}
f = function(x) exp(-x^2) / (1 + x^2)
curve(f, 0, 10, ylab = "f(x)", xlab = "x", main="graph of the function f(x)", col="purple")
```
```{r}
integrate(f,0,Inf)
```


## 1.12 (Bivariate normal). 
Construct a matrix with 10 rows and 2 columns,
containing random standard normal data:
x = matrix(rnorm(20), 10, 2)
This is a random sample of 10 observations from a standard bivariate normal
distribution. Use the apply function and your norm function from Exercise
1.10 to compute the Euclidean norms for each of these 10 observations.
```{r}
x = matrix(rnorm(20), 10, 2)
norms = apply(x, 1, norm)

print(norms)
```

## 1.13 (lunatics data). 
Obtain a five-number summary for the numeric variables in the lunatics data set (see Example 1.12). From the summary we can
get an idea about the skewness of variables by comparing the median and
the mean population. Which of the distributions are skewed, and in which
direction?
```{r}
library(BaM)
```
```{r lunatics}
data(lunatics)
lunatics
```

```{r}
summary(lunatics[2:6])
```
All of the variables are right skewed because their means are larger than their medians.


## 1.14 (Tearing factor of paper). 
The following data describe the tearing
factor of paper manufactured under different pressures during pressing. The
data is given in Hand et al. [21, Page 4]. Four sheets of paper were selected
and tested from each of the five batches manufactured.

Enter this data into an R data frame with two variables: tear factor and
pressure. Hint: it may be easiest to enter it into a spreadsheet, and then
save it as a tab or comma delimited file (.txt or .csv). There should be 20
observations after a successful import.
```{r}
tear_factor <- c(112, 119, 117, 113, 108, 99, 112, 118, 120,
      106, 102, 109, 110, 101, 99, 104, 100, 102, 96, 101)
pressure <- rep(c(35.0, 49.5, 70.0, 99.0, 140.0), each = 4)
paper_data <- data.frame(pressure, tear_factor)
paper_data
```

## 1.15 (Vectorized operations). 
We have seen two examples of vectorized
arithmetic in Example 1.1. In the conversion to Celsius, the operations involved one vector temps of length four and scalars (32 and 5/9). When we
computed differences, the operation involved two vectors temps and CT of
length four. In both cases, the arithmetic operations were applied element
by element. What would happen if two vectors of different lengths appear
together in an arithmetic expression? Try the following examples using the
colon operator : to generate a sequence of consecutive integers.

a:
```{r}
x = 1:8
n = 1:2
x+n
```
R repeats the shorter vector as many times as necessary until its length is the same as the longer vector.

b:
```{r}
n = 1:3
x+n
```
R repeats the shorter vector as many times as necessary until its length is the same as the longer vector, but in this case, throws a warning because longer vector length is not a multiple of shorter vector.